Type Classes for Mathematics in Type Theory

### 2010-12-10 papers/mscs/mscs.tex
Change citation and bibliography style in MSCS paper.

% \documentclass{mscs}
\documentclass[a4paper,10pt,runningheads]{llncs}
\usepackage{url}
\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{upgreek}
\usepackage[mathletters]{ucs}
\usepackage[utf8x]{inputenx}
\usepackage{comment}
\newcommand{\N}{\ensuremath{\mathbbm{N}}}
\newcommand{\Z}{\ensuremath{\mathbbm{Z}}}
\newcommand{\Q}{\ensuremath{\mathbbm{Q}}}

@@

The Coq proof assistant is based on the calculus of inductive
constructions~\cite{CoquandHuet,CoquandPaulin}, a dependent type theory with (co)inductive types; see~\cite{Coq,BC04}. In true Curry-Howard fashion, it is both an excessively pure, if somewhat pedantic, functional programming language with an extremely expressive type system, and a language for mathematical statements and proofs. We highlight some aspect of Coq that are of particular relevance to our development.

@@

Propositions in Coq are types~\cite{ITT,CMCP}, which themselves have types called \emph{sorts}. Coq features a distinguished sort called \lstinline|Prop| that one may choose to use as the sort for types representing propositions. The distinguishing feature of the \lstinline|Prop| sort is that terms of non-\lstinline|Prop| type may not depend on the values of inhabitants of \lstinline|Prop| types (that is, proof terms).
% As originally conceived~\cite{ITT,CMCP}, Propositions are Types and we can extract information from
% them. Often we do \emph{not} want functions to depend on proofs. In Coq, this is the paradigmatic way of working. We are forbidden to extract information from proofs of \lstinline|Prop|s. Especially with the aid of the Program machinery, this allows us to flexibly work with simply typed program and their Hoare style correctness proofs.
This regime of discrimination establishes a weak form of proof irrelevance, in that changing a proof can never affect the result of value computations. On a very practical level, this lets Coq safely erase all \lstinline|Prop| components when extracting certified programs to OCaml or Haskell.

@@

We are currently in the process of retrofitting the rationals interface into CoRN. In future work we aim to base our development of its reals on an abstract dense set, allowing us to use the efficient dyadic rationals~\cite{boldo2009combining} as a base for exact real number computation in Coq~\cite{Riemann,Oconnor:real}. The use of category theory has been important in these developments. % Eelis: not sure i understand. "has been"? it's not done.

@@

This work has been partially funded by the FORMATH project, nr.\ 243847, of the FET program within the 7th Framework program of the European Commission.
% MSCS recommends harvard, but does not specify which bst file to use.
\bibliographystyle{plain}
\bibliography{alg}
\end{document}

>>>>

% \documentclass{mscs}
\documentclass[a4paper,10pt,runningheads]{llncs}
\usepackage{url}
\usepackage{bbm}
\usepackage{amsmath}
\usepackage{harvard}
\usepackage{amssymb}
\usepackage{upgreek}
\usepackage[mathletters]{ucs}
\usepackage[utf8x]{inputenx}
\usepackage{comment}
\newcommand{\N}{\ensuremath{\mathbbm{N}}}
\newcommand{\Z}{\ensuremath{\mathbbm{Z}}}
\newcommand{\Q}{\ensuremath{\mathbbm{Q}}}

@@

The Coq proof assistant is based on the calculus of inductive
constructions~\cite{CoquandHuet,CoquandPaulin}, a dependent type theory with (co)inductive types; see~\cite{BC04,Coq}. In true Curry-Howard fashion, it is both an excessively pure, if somewhat pedantic, functional programming language with an extremely expressive type system, and a language for mathematical statements and proofs. We highlight some aspect of Coq that are of particular relevance to our development.

@@

Propositions in Coq are types~\cite{CMCP,ITT}, which themselves have types called \emph{sorts}. Coq features a distinguished sort called \lstinline|Prop| that one may choose to use as the sort for types representing propositions. The distinguishing feature of the \lstinline|Prop| sort is that terms of non-\lstinline|Prop| type may not depend on the values of inhabitants of \lstinline|Prop| types (that is, proof terms).
% As originally conceived~\cite{ITT,CMCP}, Propositions are Types and we can extract information from
% them. Often we do \emph{not} want functions to depend on proofs. In Coq, this is the paradigmatic way of working. We are forbidden to extract information from proofs of \lstinline|Prop|s. Especially with the aid of the Program machinery, this allows us to flexibly work with simply typed program and their Hoare style correctness proofs.
This regime of discrimination establishes a weak form of proof irrelevance, in that changing a proof can never affect the result of value computations. On a very practical level, this lets Coq safely erase all \lstinline|Prop| components when extracting certified programs to OCaml or Haskell.

@@

We are currently in the process of retrofitting the rationals interface into CoRN. In future work we aim to base our development of its reals on an abstract dense set, allowing us to use the efficient dyadic rationals~\cite{boldo2009combining} as a base for exact real number computation in Coq~\cite{Oconnor:real,Riemann}. The use of category theory has been important in these developments. % Eelis: not sure i understand. "has been"? it's not done.

@@

This work has been partially funded by the FORMATH project, nr.\ 243847, of the FET program within the 7th Framework program of the European Commission.
% MSCS recommends harvard, but does not specify which bst file to use.
\bibliographystyle{agsm}
\bibliography{alg}
\end{document}
